---
title: "*Which is the best bike for this route?* is not the (only) question you should be asking"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
source("main.R")
```


- Importance of bike choice; racing, chasing PRs, KOMs and sprints
- Resource of ZwiftInsider speed test data and derivatives (e.g. my data browser, ZwifterBikes)
  - Details of testing
  - Judging routes, might expect fairly aero setup to do well even on fairly climb heavy routes like Mountain Route (survey?)
  
```{r}
all_bikes[power==300][
  order(route),
  .("Frame"=frame, "Wheel"=wheel, "Time"=
      sprintf("%02.f:%02.f (+% 2s)",
              route%/%60, route%%60, as.character(route-min(route))))][
  1:10]
```

  
- But ZI tests performed using a very specific profile of rider; 300w, 182cm, 75kg
  - This doesn't match all riders on Zwift
  - Varying rider power/power-to-weight ratio should change the relative importance of aerodynamic performance and climbing performance, with slow riding placing a great emphasis on climbing performance
  - This means judging which bike is best for a given route is not just about how much climbing occurs on a route, but also needs to incorporate the profile of the rider
- To test how rider power/power-to-weight affects performance and recommended frame and wheel choice, I used a bot to test performance of a 182cm, 75kg rider riding at 300W (4W/kg) and 150W (2W/kg)
  - I did this test on Watopia's Mountain Route as a more representative route; it has some flat, some mixed difficulty of climbing and some descents
  - ZI tests were on very extreme routes, either completely flat or steep climbing (so should be less sensitive to rider variation)
  - Frame and wheel performance is independent, so wheels were tested on a Zwift Aero frame and frames tested with Zwift 32mm Carbon wheels, and frame and wheel effects used to get predicted times for each unqiue combination
    - These predicted times were verified for a random selection of bikes
  





```{r}
all_bikes[power==150&frame=="Zwift Aero"][
  order(route),
  .("Wheel"=wheel, "Time Behind (s)"=route-min(route))]

all_bikes[power==300&frame=="Zwift Aero"][
  order(route),
  .("Wheel"=wheel, "Time Behind (s)"=route-min(route))]
```


```{r}
all_bikes[power==150&frame=="Zwift Aero"][
  order(epic),
  .("Wheel"=wheel, "Time Behind (s)"=epic-min(epic))]

all_bikes[power==300&frame=="Zwift Aero"][
  order(epic),
  .("Wheel"=wheel, "Time Behind (s)"=epic-min(epic))]
```


