---
title: "Which bike is quickest up the alp?"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
source("main.R")

# survey_data <- data.table(
#   bike=c("Aethos & Lightweights", "Venge S-Works & Lightweights", "Aethos & DT Swiss 62",
#          "TRON", "Aethos & DT Swiss Disc", "Scott Addict & DT Swiss 62", 
#          "Scott Addict & Lightweights", "Venge S-Works & DT Swiss Disc"),
#   votes=c(20,2,1,72,1,9,4,1)
# )

```


The Specialized Aethos S-Works and Lightweight Meilenstein wheels, right? Nope, it's the Cadex Tri TT frame with DT Swiss ARC 1100 DiCut Disc wheels of course. No really, that's the quickest. At least, it's the quickest *if* you weigh 50 kg and can knock out 2000 Watts for 10 minutes &mdash; it's 26 seconds (>4%) quicker than the Aethos setup. Ok, I accept that not many of us are quite doing those numbers, but the point stands. **The quickest bike over a route is not *just* about the route, but also about the rider.**

Bike choice is important to get right for racers, PR chasers and KQOM hunters, where margins of mere seconds make all the difference. Go into any article about a route on Zwift, Zwift Facebook group or forum and you'll quickly find plenty of questions asking *which bike for this route?* The answers always boil down to balancing the relative needs for aerodynamic and climbing performance given the profile of the route, but few &mdash; if any &mdash; discuss the profile of the rider.



### Why does rider profile matter?

There are several forces affecting a rider, but the important one to consider here is aerodynamic drag &mdash; it increases exponentially with speed, while the other forces slowing the rider increase linearly. This means that, as speed increases, so too does the relative importance of aerodynamic performance. If a rider is going up a 5% slope at 50 Watts, aerodynamic drag is a tiny part of the forces the rider is overcoming to move forward, but if that rider is going up at 500 Watts then aerodynamic is going to be a very large component. If aerodynamic drag is a big component of the forces slowing you down, then there are potentially big gains to minimizing it. Likewise, if it's a tiny component, then there is less to be gained by reducing aerodynamic drag.



### The Experiment

With this idea in mind, I set out to perform an experiment. The aim was to test how rider profile affects bike choice. I chose to set up riding around the Mountain Route in Watopia. I chose this route for a few reasons:

1. While the Tempus Fugit and Alpe du Zwift tests shed light on performance at the extreme ends of the scale, that also means differences in relative performance are not likely to change outside of extreme conditions (e.g. 2000 Watts at 40 W/kg)
1. Mountain route is not just flat or just uphill, but more representative of routes available in Zwift
1. Mountain route contains flat sections (from start to KOM), rolling sections (Villas and Esses), a moderate pitch climb (Epic) and a steep climb (Radio Tower) and their descents, allowing me to focus in on those areas and compare performance across those route profile types

My rider, a bot &mdash; Robot Griffin &mdash; is strikingly similar to the ZwiftInisder bot. Both are 182 cm tall, weigh in at 75kg, can happily push 300 Watts (4 W/kg) for an hour, and like to ride behind a firewall to take away drafting and stop ruining leaderboards. Unlike the ZwiftInsider bot, Robot has also added in some zone two training (it's all the rage) and will happily plod along at 150 Watts (2 W/kg). 

But Robot is also time-pressed father of three and couldn't possibly test all frames and wheelsets, so he picked a few of the tried and tested standout performers. Along with the TRON, he rode the Specialized Venge S-Works and Aethos S-Works, Scott Addict RC, Canyon Aeroad 2021, Focus Izalco Max, Pinarello Dogma F and Zwift Aero frames (all on Zwift 32mm Carbon wheels) and DT Swiss ARC 1100 DiCut Disc and 62, ZIPP 808 and 454, ENVE SES 3.4 and 7.8, and Lightweight Meilenstein wheels (all on the Zwift Aero frame). Because frame and wheel performance is independent, these data can be worked to give accurate predictions of times for every combination of frame and wheel, giving times for `r all_bikes[power==300, .N]` different bikes, each tested at 150 Watts and 300 Watts.


### The Results

First up, let's look at the results at 300 Watts &mdash; the same conditions as run in the ZwiftInsider speed tests. Beforehand, just based on judgement of route profile and the ZwiftInsider speed tests, I would have picked a bike like the TRON or an all-rounder like a Scott Addict RC with ZIPP 454s. I even did a sneaky survey of the Zwift Riders Facebook group and among ten well suited bikes, the TRON was the standout favourite with  72 of the 110 votes, and the Aethos with Lightweight Meilensteins came in second with 20 votes. Surprisingly, only 1 of the 110 voters picked the fastest bike (at 300 Watts) &mdash; the `r all_bikes[power==300][order(route), frame][1]` frame and `r all_bikes[power==300][order(route), wheel][1]` wheelset!


  
  
```{r}
all_bikes[power==300][
  order(route),
  .("Frame"=frame, "Wheel"=wheel, "Time"=
      sprintf("%02.f:%02.f (+% 2s)",
              route%/%60, route%%60, as.character(route-min(route))))][
  1:10]
```

  
- But ZI tests performed using a very specific profile of rider; 300w, 182cm, 75kg
  - This doesn't match all riders on Zwift
  - Varying rider power/power-to-weight ratio should change the relative importance of aerodynamic performance and climbing performance, with slow riding placing a great emphasis on climbing performance
  - This means judging which bike is best for a given route is not just about how much climbing occurs on a route, but also needs to incorporate the profile of the rider
- To test how rider power/power-to-weight affects performance and recommended frame and wheel choice, I used a bot to test performance of a 182cm, 75kg rider riding at 300W (4W/kg) and 150W (2W/kg)
  - I did this test on Watopia's Mountain Route as a more representative route; it has some flat, some mixed difficulty of climbing and some descents
  - ZI tests were on very extreme routes, either completely flat or steep climbing (so should be less sensitive to rider variation)
  - Frame and wheel performance is independent, so wheels were tested on a Zwift Aero frame and frames tested with Zwift 32mm Carbon wheels, and frame and wheel effects used to get predicted times for each unqiue combination
    - These predicted times were verified for a random selection of bikes
  





```{r}
all_bikes[power==150&frame=="Zwift Aero"][
  order(route),
  .("Wheel"=wheel, "Time Behind (s)"=route-min(route))]

all_bikes[power==300&frame=="Zwift Aero"][
  order(route),
  .("Wheel"=wheel, "Time Behind (s)"=route-min(route))]
```


```{r}
all_bikes[power==150&frame=="Zwift Aero"][
  order(epic),
  .("Wheel"=wheel, "Time Behind (s)"=epic-min(epic))]

all_bikes[power==300&frame=="Zwift Aero"][
  order(epic),
  .("Wheel"=wheel, "Time Behind (s)"=epic-min(epic))]
```


